\chapter{Software Requirements}

\section{Introduction}

The new program is to replace WAQBANK which was developed for the SIMONA system.
The essence of the requirements is that the new code should be consistent with the WAQBANK version except for the fact that it supports new \dflowfm map-files.

WAQBANK consists of two components:

\begin{enumerate}
\item the first component to determine the location of bank lines based on simulation results and a reference bank line (e.g.~from Baseline)
\item the second component to determine the bank erosion along that line based on simulation results and user input.
\end{enumerate}

The input of WAQBANK consists of one configuration using an ini-file like syntax which includes references to SIMONA or Delft3D 4 simulation files, and various simple ASCII files containing analysis parameters varying along the river reach.
The output of WAQBANK consists of a number of simple ASCII files reporting on the erosion volumes per reach segment (e.g.~per km) and a number of files listing the new coordinates of the bank lines after erosion.
Finally, the WAQBANK program may write some data in file formats suitable for visualization in WAQVIEW and generate a number of figures.

\section{Functional Requirements} \label{Sec:FuncReq}

\begin{enumerate}
\item The results of this program must match those of WAQBANK given the same input data.
\item Users must be able to run this program in batch mode from the command line.
\item Users must be able to run the analysis based on \dflowfm results.
\item Users must be able to provide all data via an input file, similar to the ini-file like file of WAQBANK.
\item The input files must be consistent with those of WAQBANK, or aligned with open standards or the \dflowfm modeling system.
\item The output files must be consistent with those of WAQBANK, or aligned with open standards or the \dflowfm modeling system.

\item The should read relevant data directly from \dflowfm map-files similarly to WAQBANK reading data directly from SIMONA and Delft3D 4 result files.

\item A simple graphical user interface could support users in process of creating the input file.

\item It would be nice if the software would be more generally applicable than just the Dutch rivers.
\item It would be nice if the software would be able to run besides English also in Dutch.
\end{enumerate}

All requirements are addressed by \dfastbe.

\section{Non-functional requirements}

\begin{enumerate}
\item The performance of the tools must be similar to that of WAQBANK, i.e.~it must run within about an hour.
\item The software must be version controlled.
\item The software must have formal testing and support.
\item The software must run on Windows.
\item The software must be easy to distribute.
\item The software must have a user manual.
\item The software must have technical documentation.
\item The software must meet the conditions described in \href{http://publicaties.minienm.nl/documenten/rijkswaterstaat-informatievoorziening-aansluitvoorwaarden-riva-2017}{Rijkswaterstaat Informatievoorziening Aansluitvoorwaarden}.

\item The software should run on any common operating system.
\item The software should be available as open source.
\end{enumerate}

All requirements are addressed by \dfastbe although the testing has been carried out on the Windows platform only
